---
title: "Comp Shapes CogSci2025"
author: Jess Mankewitz
output:
  pdf_document: 
    toc: true
    toc_depth: 3
  html_document: default
---
# Comp Shapes CogSci2025

Data processing, analysis, and figures for the 2025 CogSci Submission titled "Compositionality Emerges from Communicative Context"

## Setup

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lmerTest)
library(here)
library(broom.mixed)
library(tidyboot)

color_map <- list('Within Trial' = "#85216BFF",'No Competitor' = "#FCAD12FF", 'Across Trial' = "#CB4149FF")

fig_write_path <- here("analysis/figures")

game_blacklist <- c("01JGZ1J4WKTCRGQS8R8214JJH2",#AI response
                   "01JGMPEGCD0EX4TXSSNEY3A670" #AI response
                   )

final_game_set <- "run_v3"

target_experiment_names <- list.files(here("data", "processed_data", final_game_set))

# condition names

old_conditions <- c('noncomp','comp-within','comp-between')
new_conditions_long <- c('No Competitor','Within Trial','Across Trial')
new_conditions_short <- c('No Comp','Comp Within','Comp Between')
```

Set personal theme

```{r}
theme_jmank <- function(base_size = 16) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      axis.line = element_line(color = "black"),
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.position = "bottom",
      legend.background = element_rect(fill = "transparent", colour = NA),
      strip.text = element_text(size = rel(0.85), face = "bold", margin = margin(5,0,5,0))
    )
}
```


## Data Importing

All data from these folders were generated using the `00.preprocessed.R` R script. 

Helper functions:

```{r}
load_data <- function(data_name, study_list) {
  source_files <- list.files(path = here("data", "processed_data", final_game_set),
             recursive = TRUE,
            pattern = paste0(data_name, ".csv$"),
             full.names = TRUE)
  # Bind the rows of only the data files in the target experiment/study list
  data_files <- source_files[Reduce("|", lapply(study_list, function(x) grepl(x, source_files)))]
  do.call(bind_rows, lapply(data_files, read.csv))
}
```

Load data :

```{r message=FALSE}
d_game <- load_data("games", target_experiment_names)
d_round <- load_data("rounds", target_experiment_names)
d_chat <- load_data("chats", target_experiment_names)
d_players <- load_data("players", target_experiment_names)
d_chat_similarity <- read_csv(here('data/processed_data/message_similarities.csv'))
```

### Relevel Condition Factor

```{r}
d_game$contextStructure_f <- factor(d_game$contextStructure, 
                                    levels=old_conditions,
                                    labels=new_conditions_long)

d_game <- d_game %>% mutate(contextStructure_f = relevel(contextStructure_f, ref = 'No Competitor'))
```

## Preregistered Exclusions

Two key exclusions (of games that were able to start successfully ie have at least 1 response):

1) Exclude games with fewer than 50% (32) responses
2) Exclude games where accuracy was less than 75%

Non-registered exclusions: Exclude games were the output is suspiciously AI-like

```{r message=FALSE, warning=FALSE}
game_stats <- d_round |> 
  left_join(d_game) |> 
  #aggregate counts across games
  group_by(gameID, contextStructure, contextStructure_f) |> 
  summarize(n_correct = sum(correct, na.rm = T),
            n_responses = sum(!is.na(response))) |>  
  filter(n_responses > 0) |> 
  mutate(non_response_rate = n_responses < 32,
         inaccuracy = n_correct/n_responses < .75,
         blacklist = gameID %in% game_blacklist,
         # exclusionary criterea
         kept_game = ! non_response_rate & ! inaccuracy & ! blacklist)

# Store the stats for each exclusion criterea for easier in-text reporting

non_response <- game_stats %>% 
  group_by(non_response_rate, inaccuracy, blacklist, kept_game) |> 
  summarise(n_games = n_distinct(gameID)) %>% 
  filter(non_response_rate)

inaccuracy <- game_stats %>% 
  group_by(non_response_rate, inaccuracy, blacklist, kept_game) |> 
  summarise(n_games = n_distinct(gameID)) %>% 
  filter(!non_response_rate, inaccuracy)

blacklist <- game_stats %>% 
  group_by(non_response_rate, inaccuracy, blacklist, kept_game) |> 
  summarise(n_games = n_distinct(gameID)) %>% 
  filter(blacklist)

final_n <- game_stats %>% 
  group_by(contextStructure, contextStructure_f, kept_game) |> 
  summarise(n_games = n_distinct(gameID)) %>% 
  pivot_wider(names_from = kept_game, values_from=n_games) %>% 
  rename("excluded" = `FALSE`, "included" = `TRUE`) %>% 
  mutate(prop_excluded = excluded/(excluded+included))
```

We're expecting 150 included games for each condition, after exclusion.

```{r}
final_n
```

Do exclusions:

```{r message=FALSE, warning=FALSE}
d_game_f <- d_game %>% left_join(game_stats %>% select(gameID, kept_game)) %>% filter(kept_game)
d_round_f <- d_round %>% left_join(game_stats %>% select(gameID, kept_game)) %>% filter(kept_game)
d_chat_f <- d_chat %>% left_join(d_round_f %>% select(roundID, kept_game, gameID)) %>% filter(kept_game)
d_players_f <- d_players %>% left_join(game_stats %>% select(gameID, kept_game)) %>% filter(kept_game)
d_chat_similarity_f <- d_chat_similarity %>% left_join(game_stats %>% select(gameID, kept_game)) %>% filter(kept_game)
```

### Paper Text

Dyads were excluded from analysis (but still compensated) if they were missing more than 32 (50%) trials (n=`r sum(non_response$n_games)`) or failed to identify the correct target more than 75% of the time (n=`r inaccuracy$n_games`). Two additional dyads were excluded because their messages displayed clear markers of AI-generated text (e.g., unusually long, instructional descriptions like "The black box surrounding it suggests it is a target object...Let me know if you need more insights!"). The final sample includes `r sum(final_n$included)` dyads across three conditions (n=`r final_n %>% filter(contextStructure == "noncomp") %>% pull(included)` dyads in the no-competitor condition, n=`r final_n %>% filter(contextStructure == "comp-within") %>% pull(included)` in the within-trial competitor condition, and n=`r final_n %>% filter(contextStructure == "comp-between") %>% pull(included)` in the across-trial competitor condition). Participants received a base payment of \$8.25 plus a \$0.03 bonus for each correct response.

## Reference Game Replications

### Accuracy over time

First, we examine how the dyads perform over the course of the game, looking at the accuracy over time across games in different conditions.

#### Data

```{r}
d_rounds_game <- d_round_f %>% 
  filter(kept_game)  |> 
  left_join(d_game_f) %>% 
  mutate(index_scaled = scale(index))

d_accuracy <- d_round_f %>% 
  filter(kept_game)  |> 
  left_join(d_game_f) |> 
  filter(!is.na(response)) |> 
  group_by(index, contextStructure_f) |> 
  summarize(n_games = n_distinct(gameID),
            n_correct = sum(correct),
            prop_correct = n_correct/n_games)
```

#### Model

```{r}
# Fit logistic mixed effects model 
acc_model <- glmer(correct ~ contextStructure_f * index_scaled + 
                   (1 + index_scaled | gameID) + (1 + index_scaled | target),
                   family = binomial,
                   control = glmerControl(optimizer = "bobyqa"),
                   data = d_rounds_game)

# Get model summary
summary(acc_model)
```

#### Plot

Summarized across games

```{r message=FALSE, warning=FALSE}
d_acc <- d_round_f %>% 
  filter(kept_game)  |> 
  left_join(d_game_f) |> 
  filter(!is.na(response)) |> 
  group_by(index, contextStructure_f) |> 
  summarize(n_games = n_distinct(gameID),
            n_correct = sum(correct),
            prop_correct = n_correct/n_games)

p_acc <- d_acc %>% 
  ggplot(aes(x = index, y = prop_correct, color = contextStructure_f)) + 
  geom_vline(xintercept = c(16, 32, 48), linetype = "dashed") +
  geom_point(alpha = .3) + 
  theme_minimal() + 
  geom_smooth(aes(fill = contextStructure_f), se=F, method = "glm", 
              method.args = list(family = "binomial")) +
  labs(title = "Dyads Quickly Become Accurate",
       x = "Trial Number", 
       y = "Proportion of Correct Responses", 
       color = "Subshape Competitor",
       fill = "Subshape Competitor") + 
  scale_color_manual(values = color_map) + 
  scale_fill_manual(values = color_map) + 
  theme_jmank() + 
  guides(color=guide_legend(title.position="top"),
         fill=guide_legend(title.position="top"))

p_acc
ggsave(filename = here(fig_write_path, "prop_correct.png"), 
       width = 3.5, height = 3)
```

### Word Count

Count the number of words in each utterance.

```{r message=FALSE}
d_chat_cleaned <- d_chat_f %>% 
  full_join(d_round_f) |> 
  left_join(d_game_f) |> 
  filter(kept_game) %>% 
 filter(director_msg) %>% 
  filter(!chit_chat | is.na(chit_chat)) %>% #filter out chit-chat
  mutate(text = gsub('[[:punct:] ]+',' ',text), #stripping some punctuation
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1)

d_chat_word_counts <-d_chat_cleaned %>% 
  group_by(gameID, index, repNum, playerID, target, contextStructure_f) %>%
  summarize(text = paste0(text, collapse = ', '),
            total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars))
```

#### Paper Text
Word counts:


The processed dataset contains `r nrow(d_chat_word_counts)` unique referring expressions. All messages were cleaned by removing punctuation, removing trailing white-space, and converting to lowercase.

#### Model

```{r}
word_count_model <- glmer(total_num_words ~ contextStructure_f * poly(index, 2) +
                         (1 + poly(index, 1) | gameID),  
                         family = poisson,
                         control = glmerControl(optimizer = "bobyqa"),
                         data = d_chat_word_counts)
summary(word_count_model)
```
#### Model Paper Text

Another hallmark of convention formation is the reduction of referring expression length over time. We modeled word counts using a Poisson mixed-effects regression with condition, trial number (modeled with orthogonal polynomials to capture non-linear trends), and their interaction as fixed effects, plus by-dyad random intercepts and slopes for trial number. The analysis revealed a significant reduction in description length over time (beta = -34.68, SE = 2.11, p < .001; Fig. 3b). Notably, conditions differed in their overall verbosity - participants in the within-trial competitor condition used longer descriptions than baseline (beta = 0.15, SE = 0.06, p = .01), while the across-trial competitor condition used shorter ones (beta = -0.34, SE = 0.06, p < .001). However, this reduction in length alone does not necessarily indicate successful convention formation - participants may simply become more comfortable with the task over time without establishing stable referring expressions. To test whether true conventions emerged, we next examine how descriptions for individual shapes evolved over repeated references.

#### Plot

```{r}
d_word_counts <- d_chat_word_counts %>% 
  group_by(index, contextStructure_f) %>% 
  summarize(mean_n_words = mean(total_num_words, na.rm = T),
            sd_n_words = sd(total_num_words, na.rm = T))

p_word_counts <- d_word_counts %>% 
  ggplot(aes(x = index + 1, y = mean_n_words, color = contextStructure_f)) +
  geom_vline(xintercept = c(16, 32, 48), linetype = "dashed") +
  geom_point(alpha = .3) + 
  geom_smooth(aes(fill = contextStructure_f), 
              method = "glm", se=F, 
              formula = y ~ x + I(x^2)) +
  labs(x = "Trial Number", 
       y = "Mean Number of Words", 
       title = "Directors Use Fewer Words", 
      color = "Subshape Competitor",
       fill = "Subshape Competitor") + 
  scale_color_manual(values = color_map) + 
  scale_fill_manual(values = color_map) + 
  theme_jmank() +
  guides(color = "none", fill="none")

p_word_counts
```


### Across Block Transition Utterance Similarity

Here, we're looking at how similar the descriptions are for a single tangram from one round to the next. 

```{r message=FALSE}
d_chat_similarity_full <- d_chat_similarity_f |> 
  left_join(d_round |> select("roundID1"="roundID","target1"="target", "index1" = "index", "repNum1" = "repNum")) |> 
  left_join(d_round |> select("roundID2"="roundID","target2"="target", "index2" = "index", "repNum2" = "repNum")) |> 
  left_join(d_game |> select("gameID", "contextStructure_f", "rotation"))
```


```{r message=FALSE}
sim_by_tangram <- d_chat_similarity_full |> 
   # chats for the same target
  filter(target1 == target2) |> 
    rowwise() |> 
  # create transition labels between blocks
    mutate(rep_trans = paste(min(repNum1+1, repNum2+1), 
                             max(repNum1+1, repNum2+1), sep = "->")) |> 
    ungroup() |> 
    filter(rep_trans %in% c("0->1", "1->2","2->3", "3->4")) |> 
  # order the factor for proper plotting
  mutate(rep_trans = factor(rep_trans, 
                           levels = c("0->1", "1->2", "2->3", "3->4"))) %>% 
  group_by(gameID, rep_trans, target1, contextStructure_f) |> 
  summarize(mean_sim = mean(similarity, rm.na = T),
            sd_sim = sd(similarity))

sim_by_game <- sim_by_tangram |> 
  group_by(gameID, rep_trans, contextStructure_f) |> 
  summarize(mean_sim = mean(mean_sim, rm.na = T),
            sd_sim = sd(mean_sim))

sim_across_game <- sim_by_game |> 
  group_by(rep_trans, contextStructure_f) |> 
  summarize(mean_sim = mean(mean_sim, rm.na = T),
            sd_sim = sd(mean_sim))
```


#### Model

```{r}
convention_model <- lmer(mean_sim ~ rep_trans * contextStructure_f + 
                        (1 | gameID),
                        data = sim_by_tangram)

summary(convention_model)
```

#### Paper Text

3.1.3 Description Stability

To confirm that participants formed true conventions rather than simply producing shorter descriptions due to task familiarity, we examined how descriptions for the same shape evolved across repeated references. We measured stability by computing cosine similarities between SBERT embeddings of descriptions for the same tangram across consecutive blocks, and modeled these similarities using a linear mixed-effects regression with condition, block transition, and their interaction as fixed effects, plus random intercepts for dyad.

The analysis revealed increasing similarity across blocks in all conditions, with descriptions becoming more similar between blocks 2-3 (beta = 0.07, SE = 0.007, p < .001) and blocks 3-4 (beta = 0.14, SE = 0.007, p < .001; Fig. 3c). Notably, this conventionalization was stronger in both compositional conditions compared to the baseline (block 3-4 increases: Within beta = 0.04, Between beta = 0.07, both p < .001). This suggests that the presence of shared visual components supported more stable convention formation, regardless of how these components were distributed in the communicative context. Having established successful convention formation across conditions, we next examine how the structure of these conventions differed based on communicative context.

#### Plot


```{r warning=FALSE}
d_sim_model <- sim_by_game %>% 
  group_by(contextStructure_f, rep_trans) %>% 
  tidyboot_mean(column = mean_sim)

p_sim_trans <- d_sim_model %>% 
  ggplot(aes(x = rep_trans, y = empirical_stat, color = contextStructure_f)) + 
  geom_jitter(data = sim_by_game, 
              aes(x = rep_trans, y = mean_sim), alpha = .075, 
                position = position_dodge(width = .2))+
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper, 
                    fill = contextStructure_f, 
                    group = contextStructure_f), 
                color = NA, alpha = .3)+
  geom_point(stat = 'identity') +
  geom_line(aes(group = contextStructure_f), size = 1) +
  labs(x = "Block Transition",
       y = "Pairwise Similarity",
       title = "Similarity of Shape Descriptions",
       subtitle = "In Consecutive Blocks",
      color = "Subshape Competitor",
       fill = "Subshape Competitor") + 
  scale_color_manual(values = unlist(color_map)) + 
  scale_fill_manual(values = color_map) + 
  ylim(.45, .9) + 
  guides(color=guide_legend(title.position="top"),
         fill=guide_legend(title.position="top")) +
  theme_jmank() +
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        legend.justification = "left")

p_sim_trans
```